Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-02-14 alloscore_manuscript/alloscore-application-shorter.tex
various edits, rephrasing

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Fri Feb  9 13:33:13 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
CU-select & 669 & 133 & 774 & 326 \\ 
  per-capita & 865 & - & 1029 & 366 \\ 
  COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
  USC-SI\_kJalpha & 995 & 91 & 1216 & 1097 \\ 
  JHUAPL-Gecko & 1034 & 164 & 1141 & 418 \\ 
  MUNI-ARIMA & 1084 & 169 & 1248 & 440 \\ 
  COVIDhub-trained\_ensemble & 1089 & 169 & 1271 & 823 \\ 
  COVIDhub-baseline & 1175 & 170 & 1317 & 535 \\ 
  JHUAPL-Bucky & 1358 & 102 & 1566 & 1214 \\ 
  JHUAPL-SLPHospEns & 1540 & 129 & 1604 & 1102 \\ 
  UVA-Ensemble & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The x-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the per-capital allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
\end{figure}

@@

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Fri Feb  9 13:33:17 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
model & MAS & MWIS \\ 
  \hline
COVIDhub-ensemble & 389 & 70 \\ 
  per-capita & 464 & - \\ 
  COVIDhub-trained\_ensemble & 483 & 87 \\ 
  CU-select & 502 & 81 \\ 
  JHUAPL-SLPHospEns & 526 & 67 \\ 
  COVIDhub-baseline & 594 & 93 \\ 
  JHUAPL-Bucky & 643 & 112 \\ 
  MUNI-ARIMA & 707 & 116 \\ 
  JHUAPL-Gecko & 929 & 110 \\ 
  USC-SI\_kJalpha & 1473 & 155 \\ 
   \hline
\end{tabular}
\caption{Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks. For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks. Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top. WIS is not defined for the per-capita model as this approach only provides a mechanism to allocate resources, it does not provide probabilistic forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The x- and y-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.}\label{fig:metrics-correlation}
\end{figure}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/multi-k-1} \caption[Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20]{Allocation scores (AS) across different resource constraints (K) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of K between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS). The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (light orange triangles) uses a uniform weight for all possible values of K. Note that the AS used in earlier sections of the application uses the single fixed value of K=15,000.  Panel C shows how either method of computing the IAS (y-axis) is correlated with AS at K=15,000 (x-axis). Every point represents the AS and IAS for one model. The IAS centered at 15,000 are more closely correlated with the AS values.}\label{fig:multi-k}
\end{figure}

>>>>

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Wed Feb 14 01:08:30 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrrrr}
  \hline
Model & AS & MWIS & IAS centered at 15k & IAS uniform \\ 
  \hline
CU-select & 669 & 133 & 774 & 326 \\ 
  per-capita & 865 & - & 1029 & 366 \\ 
  COVIDhub-ensemble & 873 & 159 & 1067 & 438 \\ 
  USC-SI\_kJalpha & 995 & 91 & 1216 & 1097 \\ 
  JHUAPL-Gecko & 1034 & 164 & 1141 & 418 \\ 
  MUNI-ARIMA & 1084 & 169 & 1248 & 440 \\ 
  COVIDhub-trained\_ensemble & 1089 & 169 & 1271 & 823 \\ 
  COVIDhub-baseline & 1175 & 170 & 1317 & 535 \\ 
  JHUAPL-Bucky & 1358 & 102 & 1566 & 1214 \\ 
  JHUAPL-SLPHospEns & 1540 & 129 & 1604 & 1102 \\ 
  UVA-Ensemble & 2469 & 213 & 2635 & 2494 \\ 
   \hline
\end{tabular}
\caption{For one illustrative week, a comparison of allocation scores (AS), mean weighted interval scores (MWIS), and two varieties of Integrated Allocation Scores (IAS). All metrics are shown for 10 models that made forecasts of hospital admissions for 2022-01-03. Results are sorted by AS. For all metrics, lower scores indicate better accuracy.} 
\label{tab:multi-k-scores}
\end{table}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-over-time-1} \caption[Hospital admissions and evaluation metrics over time]{Hospital admissions and evaluation metrics over time. Panel A shows the number of hospital admissions in the US as a whole due to COVID-19 on a sequence of 15 Mondays from December 2021 through March 2022. These are the dates for which forecasts were made and evaluated. A horizontal dashed line at 15,000 shows the hypothetical resource constraint $K$. Panel B shows allocation scores (AS) for each model's 14 day-ahead forecast, across all US states. The $x$-axis corresponds to the date of the observation that a model's prediction was targeting (e.g., the date the forecast was made plus the forecast horizon). AS typically are high when the observed value is near to the constraint, which occurs during the last Monday in December (on the way up) and the last Monday in January (on the way down). In Panel B, the per-capita allocation is drawn in a heavier solid line. Panel C shows the MWIS metric across weeks, averaged across all states. MWIS values tend to scale with the observed and predicted values, and the peak MWIS values happen around and just after the peak of the Omicron wave.}\label{fig:metrics-over-time}
\end{figure}

@@

% latex table generated in R 4.3.1 by xtable 1.8-4 package
% Wed Feb 14 01:08:35 2024
\begin{table}[ht]
\centering
\begin{tabular}{lrr}
  \hline
model & MAS & MWIS \\ 
  \hline
COVIDhub-ensemble & 389 & 70 \\ 
  per-capita & 464 & - \\ 
  COVIDhub-trained\_ensemble & 483 & 87 \\ 
  CU-select & 502 & 81 \\ 
  JHUAPL-SLPHospEns & 526 & 67 \\ 
  COVIDhub-baseline & 594 & 93 \\ 
  JHUAPL-Bucky & 643 & 112 \\ 
  MUNI-ARIMA & 707 & 116 \\ 
  JHUAPL-Gecko & 929 & 110 \\ 
  USC-SI\_kJalpha & 1473 & 155 \\ 
   \hline
\end{tabular}
\caption{Mean weighted interval scores (MWIS) and mean allocation scores (MAS) by model across 13 weeks. For the nine models that submitted forecasts for every week from 2021-11-29 through 2022-02-21, plus a per-capita allocation, these results show the average performance of each model across all weeks. Lower scores indicate better performance, and models are sorted with the most accurate allocations at the top. WIS is not defined for the per-capita model as this approach only provides a mechanism to allocate resources, it does not provide probabilistic forecasts.} 
\label{tab:multi-week-performance-summary}
\end{table}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/metrics-correlation-1} \caption[Association of standardized ranks for MWIS and allocation score by model and week]{Association of standardized ranks for MWIS and allocation score by model and week. Each facet of the plot corresponds to one model. Within each facet, each point corresponds to a week. The $x$- and $y$-values correspond to the MWIS standardized rank and the allocation score standardized rank for that week. Points corresponding to earlier dates have darker shading. The size of the point corresponds to the observed value on the date for which the prediction was made. Models show different degrees of association between the two metrics.}\label{fig:metrics-correlation}
\end{figure}

@@

\begin{knitrout}
\definecolor{shadecolor}{rgb}{0.969, 0.969, 0.969}\color{fgcolor}\begin{figure}
\includegraphics[width=\maxwidth]{figure/multi-k-1} \caption[Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20]{Allocation scores (AS) across different resource constraints ($K$) for 10 models that made forecasts on 2021-12-20. Panel A shows, for each model, the AS for values of $K$ between 200 and 60,000 at increments of 200. The AS show a sharp peak just under 20,000, near the eventually observed number of hospitalizations. Panel B shows two possible weighting functions for the Integrated Allocation Score (IAS). The first (dark blue circles) computes weights proportional to a normal distribution centered at 15,000 (solid vertical gray line) with a standard deviation of 3,000, and truncated to be between 5,000 and 25,000. The second (light orange triangles) uses a uniform weight for all possible values of $K$. Note that the AS used in earlier sections of the application uses the single fixed value of $K=15,000$.  Panel C shows how either method of computing the IAS ($y$-axis) is correlated with AS at $K=15,000$ ($x$-axis). Every point represents the AS and IAS for one model. The IAS centered at 15,000 are more closely correlated with the AS values.}\label{fig:multi-k}
\end{figure}
